```{r echo=FALSE, eval=FALSE}
library(rmarkdown)
render('create-synthetic.Rmd')

```

```{r message=FALSE}
library(knitr)
library(MASS)
library(dplyr)
library(reshape2)
library(ggplot2)

```

```{r config, echo=FALSE}
opts_chunk$set(fig.width=12, fig.height=8)

```

## Synthetic data

Parameters for data generation.
```{r}
.G <- 10  # Number of predictors (features)
.R <- 6   # Number of targets (outputs)
.T <- 21  # How many time points to regress over
tmin <- 0  # Start of domain
tmax <- 10  # End of domain
num.obs <- 7  # How many data to condition on?
sigma.noise <- .1  # How much noise
mean.num.betas <- 4  # How many non-zero betas on average
sigma.beta <- 1  # S.d. of non-zero betas
num.Y.to.condition.on <- 7  # How many true Y to condition on to estimate mean and variance
set.seed(1)

```

Sample data to condition on to create some interesting $X$.
```{r}
# Evenly spaced points
linspace <- function(xmin, xmax, num.samples) {
    (0:(num.samples-1)) * (xmax-xmin) / (num.samples-1) + xmin
}
# Set up the observed time points
sample.times <- data.frame(time=linspace(tmin, tmax, num.obs))
sample.times$tau <- 1:nrow(sample.times)
# Create a squared exponential covariance function (RBF)
create.SE.cov.fn <- function(l=1) {
    function(t1, t2) {
        d <- outer(t1, t2, "-")
        exp(-(d/l)**2/2)
    }
}
# Multiply a covariance function by a constant
multiply.cov.fn <- function(k, sigma2) {
    function(t1, t2) sigma2 * k(t1, t2)
}
# Add diagonal noise term to a covariance matrix
add.noise <- function(K, sigma) {
    K + diag(sigma**2, nrow(K), ncol(K))
}
# Calculate the covariance
SE.cov.fn <- create.SE.cov.fn(l=3)
K.SE <- SE.cov.fn(sample.times$time, sample.times$time)
# Sample the data
obs <- mvrnorm(n=.G, mu=rep(0, nrow(sample.times)), Sigma=K.SE)

```

Plot the samples.
```{r}
# Melt into format for plotting
obs.l <- melt(obs, varnames=c("g", "tau"), value.name="y")
# Plot
(ggplot(obs.l %>% left_join(sample.times),
        aes(x=time, y=y))
    + geom_line()
    + facet_wrap(~ g))

```

Condition on the samples to get posterior for $X$.
```{r}
times <- data.frame(time=linspace(tmin, tmax, .T))
times$t <- 1:nrow(times)
# Calculate the posterior mean, variance (full) and log marginal likelihood
calc.posterior <- function(X, y, k, sigma.noise, xstar) {
    within(list(), {
        X <- X
        y <- y
        sigma.noise <- sigma.noise
        xstar <- xstar
        K <- k(X, X)
        Kstar <- k(xstar, X)
        Kstarstar <- k(xstar, xstar)
        L <- chol(add.noise(K, sigma.noise))
        alpha <- solve(L, solve(t(L), y))
        mu <- as.vector(Kstar %*% alpha)
        log.marg.like <- (
            - t(y) %*% alpha / 2
            - sum(diag(L))
            - length(y)/2*log(2*pi))
        V <- Kstarstar - Kstar %*% solve(L, solve(t(L), t(Kstar)))
    })
}
# Calculate the posteriors for each predictor
X.posteriors <- lapply(
    1:.G,
    function(r) calc.posterior(sample.times$time,
                               obs[r,],
                               SE.cov.fn,
                               sigma.noise,
                               times$time))
# Gather elements of the posterior list into a matrix/array
gather <- function(posteriors, name) {
    aperm(vapply(posteriors,
                 function(posterior) posterior[[name]],
                 FUN.VALUE=posteriors[[1]][[name]]))
}
# Gather the posterior means for the X
mu.x <- gather(X.posteriors, "mu")
Sigma.x <- gather(X.posteriors, "V")

```

Sample $X$, the 'true' values for each predictor.
```{r}
X <- t(sapply(
    X.posteriors,
    function(posterior) mvrnorm(mu=posterior$mu, Sigma=posterior$V)))

```

Plot the observed posteriors for $X$ and the 'true' values.
```{r}
X.mu <- mu.x %>% melt(varnames = c("g", "t"), value.name = "mu")
X.Sigma <- (
    Sigma.x
    %>% melt(varnames = c("g", "t", "t2"), value.name = "V")
    %>% filter(t == t2))
X.l <- X %>% melt(varnames = c("g", "t"), value.name = "X")
(ggplot(X.mu %>% left_join(X.Sigma) %>% left_join(times), aes(x=time, y=mu))
    + geom_line()
    + geom_ribbon(aes(ymin=mu-2*sqrt(V), ymax=mu+2*sqrt(V)), alpha=.5)
    + geom_point(data=X.l %>% left_join(times),
                 aes(x=time, y=X),
                 size=3,
                 alpha=.4)
    + facet_wrap(~ g))

```

Sample $\beta$ for each predictor.
```{r}
sample.betas <- function(.G, mean.num, sigma.beta=1) {
    stopifnot(mean.num >= 1)
    how.many <- min(1 + rpois(1, mean.num - 1), .G)
    idxs <- sample(1:.G, how.many)
    betas <- rep(0, .G)
    betas[idxs] <- rnorm(n=how.many, mean=0, sd=sigma.beta)
    betas
}
betas <- sapply(1:.R, function(r) sample.betas(.G, mean.num.betas, sigma.beta))

```

Calculate $Y = \beta' X$, the 'true' targets.
```{r}
Y <- t(betas) %*% X

```

Select a few time points for each $y$ and condition on them to get
an observed posterior for $Y$.
```{r}
condition.Y <- function(r) {
    condition.idxs <- sample(times$t, num.Y.to.condition.on)
    posterior <- calc.posterior(times$time[condition.idxs],
                                Y[r, condition.idxs],
                                multiply.cov.fn(SE.cov.fn, mean.num.betas**2),
                                sigma.noise*mean.num.betas,
                                times$time)
    posterior$conditioned <- rep(FALSE, ncol(Y))
    posterior$conditioned[condition.idxs] <- TRUE
    posterior
}
Y.posteriors <- lapply(1:.R, condition.Y)
# Gather the posterior means for the Y
mu.y <- gather(Y.posteriors, "mu")
Sigma.y <- gather(Y.posteriors, "V")
conditioned.y <- gather(Y.posteriors, "conditioned")

```

Plot the observed posteriors for Y with the 'true' Y.
```{r}
Y.l <- Y %>% melt(varnames = c("r", "t"), value.name = "Y")
Y.conditioned.l <- conditioned.y %>% melt(varnames = c("r", "t"),
                                          value.name = "conditioned")
Y.mu <- mu.y %>% melt(varnames = c("r", "t"), value.name = "mu")
Y.Sigma <- (
    Sigma.y
    %>% melt(varnames = c("r", "t", "t2"), value.name = "V")
    %>% filter(t == t2))
(ggplot(Y.mu %>% left_join(Y.Sigma) %>% left_join(times), aes(x=time, y=mu))
    + geom_line()
    + geom_ribbon(aes(ymin=mu-2*sqrt(V), ymax=mu+2*sqrt(V)), alpha=.5)
    + geom_point(data=Y.l %>% left_join(times) %>% left_join(Y.conditioned.l),
                 aes(x=time, y=Y, color=conditioned),
                 size=3,
                 alpha=.8)
    + facet_wrap(~ r))

```

Create a list that can be used to pass data to the Stan model.
```{r}
stan.data <- list(
    T=.T,
    G=.G,
    R=.R,
    lambda=1,
    mu_x=mu.x,
    Sigma_x=Sigma.x,
    mu_y=mu.y,
    Sigma_y=Sigma.y)

```

Save the workspace.
```{r}
save.image('create-synthetic.RData')

```

R version and session information.
```{r}
sessionInfo()

```
